Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-08-28 supplement.tex
Last modifs
Work on comments of reviewer R1---leave some open todo notes

Crucially, one can show that, as long as the network the arguments $\dep(f)$ of a function $f$ are materialized by a clique within the network $\network$, then there exists at least one node in the tree decomposition that contains $\dep(f)$, thus allowing the evaluation of $f$.
\begin{lemma}\label{lem:cliques}
Let $G=(V,E)$ be an undirected graph, and  $T,\chi$ be a tree decomposition for $G$. For each clique $C\subset V$, there exists a node $n\in T$ such that $C\subset\chi(v)$.
\end{lemma}
\begin{proof}
  \newcommand{\TransVars}[1]{{\rm tverts}(#1)}
  \newcommand{\Children}[1]{{\rm Children}(#1)}
  For the sake of simplicity, let us consider $T$ as rooted on an arbitrary node, inducing an orientation, and denote by $\TransVars{n}\in V$ the set of vertices found in a node $n$ or its descendants, namely
    $$\TransVars{n} = \chi(n) \bigcup_{n'\in\Children{n}} \TransVars{n'}.$$

@@

RNA design introduces constraints due to the requirement of canonical
(aka complementary) base pairing, which can be exploited in a
particularly simple and effective way to reduce the complexity.  
As previously noted by \citet{Flamm2001}, the base pair complementarity induces a
bi-partition of each connected component within the base pair
dependency graph, such that the nucleotides assigned to the two set of
nodes in the partition are restricted to values in $\{\Ab,\Gb\}$ and
$\{\Cb,\Ub\}$ respectively. We call a partial sequence
\Def{cc-valid}, iff its determined positions are consistent with such a separation
for all determined positions of the same connected component.

@@

One can now modify Alg.~\ref{alg:pf}, on a tree decomposition
$(T,\chi)$, such that the message values
$\evalfor{\Message{u}{v}}{\val}$ are only computed for cc-valid
partial sequences $\val\in\partseqs(\separator{u}{v})$. Moreover, the
loop over $\val'\in\partseqs(\difference{u}{v})$ is restricted, such
that $\substitute{\val}{\val'}$ are cc-valid. Analogous restrictions are 
then implemented in the sampling algorithm Alg.~\ref{alg:sampling}.

@@

Finally, we remark that even stronger time savings could be possible
in practice, since cc-valid partial sequences can still violate
complementarity constraints, e.g. by assigning C and A to positions in
different sets of a partition, thus satisfying the bi-partition
constraints, where the positions base pair directly, rendering the
partial sequence invalid. Moreover, applications of the sampling
framework, can introduce additional constraints that further reduce
the number of valid partial subsequences. However, exploiting all such
constraints, in a complete and general way, would likely cause significant
implementation overhead, while not significantly improving
the asymptotic complexity.

@@

\section{Contribution of MultiDefect terms to the improvement}
\label{sec:improvements}
\begin{figure}[h!]
\includegraphics[width=.48\textwidth]{Figs/statistics-Term1}
\includegraphics[width=.48\textwidth]{Figs/statistics-Term2}
\caption{Break down of average MultiDefect values into their two components.}
\end{figure}

>>>>

Crucially, one can show that, as long as the arguments $\dep(f)$ of a function $f$ form a clique within the network $\network$, then there exists at least one node in the tree decomposition that contains $\dep(f)$, thus allowing the evaluation of $f$.
\begin{lemma}\label{lem:cliques}
\todo[inline]{R1: guess it should read "node v" instead of "n"!? please update the proof accordingly!}
Let $G=(V,E)$ be an undirected graph, and  $(T,\chi)$ be a tree decomposition for $G$. For each clique $C\subset V$, there exists a node $n\in T$ such that $C\subset\chi(v)$.
\end{lemma}
\begin{proof}
  \newcommand{\TransVars}[1]{{\rm tverts}(#1)}
  \newcommand{\Children}[1]{{\rm Children}(#1)}
  For the sake of simplicity, let us consider $T$ as rooted on an arbitrary node, inducing an orientation, and denote by $\TransVars{n}\in V$ the set of vertices found in a node $n$ or its descendants, namely
    $$\TransVars{n} = \chi(n) \bigcup_{n'\in\Children{n}} \TransVars{n'}.$$

@@

The given structures in the RNA design problem impose constraints on
the sequences for ensuring canonical (aka complementary)
base pairing. These constraints can be exploited in a particularly
simple and effective way to reduce the complexity.  As previously
noted by \citet{Flamm2001}, the base pair complementarity induces a
bi-partition of each connected component within the base pair
dependency graph, such that the nucleotides assigned to the two set of
nodes in the partition are restricted to values in $\{\Ab,\Gb\}$ and
$\{\Cb,\Ub\}$ respectively. We call a partial sequence \Def{cc-valid},
iff its determined positions are consistent with such a separation for
all determined positions of the same connected component.

@@

One can now modify Alg.~\ref{alg:pf}, on a tree decomposition
$(T,\chi)$, such that the values of messages
$\Message{u}{v}}$ are computed only for cc-valid
partial sequences $\val\in\partseqs(\separator{u}{v})$. Moreover, the
loop over $\val'\in\partseqs(\difference{u}{v})$ is restricted, such
that $\substitute{\val}{\val'}$ are cc-valid. Analogous restrictions are 
then implemented in the sampling algorithm Alg.~\ref{alg:sampling}.

@@

Finally, we remark that even stronger time savings could be possible
in practice, since cc-valid partial sequences can still violate
complementarity constraints, e.g. by assigning C and A to positions in
different sets of a partition, thus satisfying the bi-partition
constraints, where the positions base pair directly, rendering the
partial sequence invalid. Moreover, applications of the sampling
framework can introduce additional constraints that further reduce
the number of valid partial subsequences. However, exploiting all such
constraints, in a complete and general way, would likely cause significant
implementation overhead, while not significantly improving
the asymptotic complexity.

@@

\section{Contribution of terms to the MultiDefect score}
\label{sec:improvements}
\begin{figure}[h!]
\includegraphics[width=.48\textwidth]{Figs/statistics-Term1}
\includegraphics[width=.48\textwidth]{Figs/statistics-Term2}
\caption{Break-down of average MultiDefect values into their two components.}
\end{figure}
